DeepSphere: a graph-based spherical CNN

### 2019-11-15 rebuttal.md

Ref.
1. Semi-Supervised Classification with Graph Convolutional Networks. ICLR'17.
2. Covariant Compositional Networks For Learning Graphs. ICLR'18

### Answer

Thank you for your interest and questions.

1. First of all, the operator needs to be a Laplacian (why?), excluding the single-parameter variant of GCN. The two-parameter variant is ChebNet with a polynomial order of K=1. The method scales linearly with K, which is an hyper-parameter to be set (analogous to the kernel size in classical CNNs).
TODO(mdeff): check GCN paper for the variants
TODO(nati version):
1. Practically, we believe that any graph neural network variant may work and some might even slightly beat ChebyNet.
However, our equivariance theory utilizes the graph Laplacian as a base building block.
Hence, we do not know the theoretical properties of other GCNs such as the single parameter variant from T.K. Kipf.
Note that the two-parameter variant from T.K. Kipf is ChebNet with a polynomial order of K=1.
Eventually, the method we use scales linearly with K, which is an hyper-parameter to be set (analogous to the kernel size in classical CNNs).
2. Regarding the sentence 
"It is, however, possible to construct a CNN in which the activations transform in a predictable and reversible way,",
we read in the same paper, two sensences above:
"It can be shown that one cannot construct a CNN for images that behaves in a quasi-invariant way with respect to both translations and rotations unless every filter is directionless.".
In the case of our contribution, the graph convolution we use can be considered as directionless (isotropic convolution kernel). Hence, we do not need any special activation functions to be invariant.

Note also that on the sphere and the plane the meaning of rotation is different. 
The plane has 2 axes for translation and 1 axis for rotation, while the sphere has only 3 axes of rotation. 
In the context of this question, we assume that the axes of rotation on the sphere correspond to both rotation and translation on the plane.

### 2019-11-15 main.tex
make it a standalone script
fix notation laplacian
edit on text related to equivariance figure

% 1) summarize the work of Khasanova
Considering the equiangular sampling and graphs where each vertex is connected to 4 neighbors (north, south, east, west), \citet{khasanova2017sphericalcnn} designed a weighting scheme to minimize \eqref{eq:mean equivariance error} for longitudinal and latitudinal rotations\footnote{Equivariance to longitudinal rotation is essentially given by the equiangular sampling.}.
% because longitudinal rotation can be discrete
Their solution gives weights inversely proportional to Euclidean distances:
\begin{equation} \label{eqn:weights:khasanova}
	w_{ij} = \frac{1}{\norm{x_i-x_j}}.
\end{equation}
%leading to a high mean equivariance error if the mean is computed over rotations in SO(3).
Note that while the resulting convolution is not equivariant to the whole of $SO(3)$, that is enough for omnidirectional imaging because, as gravity consistently orients the sphere, objects only rotate longitudinally or latitudinally.

% 2) explain our approach
To achieve equivariance to all rotations, we take inspiration from \citet{belkin2005towards}.
They prove that for a \emph{random uniform sampling}, the graph Laplacian $\b{L}$ built from weights
\begin{equation} \label{eqn:weights:belkin}
	w_{ij} = e^{- \frac{1}{4t} \norm{x_i - x_j}^2}
\end{equation}
converges to the Laplace-Beltrami operator $\Delta_{\mathbb{S}^2}$ as the number of samples grows to infinity.
This result is a good starting point as $\Delta_{\mathbb{S}^2}$ commutes with rotation, i.e., $\Delta_{\mathbb{S}^2}\Rg = \Rg\Delta_{\mathbb{S}^2}$.
% and (ii) is diagonalized by the spherical harmonics.
While the weighting scheme is full (i.e., every vertex is connected to every vertex), most weights are small due to the exponential.
We hence make an approximation to limit the cost of the convolution \eqnref{graph_conv} by only considering the $k$ nearest neighbors to each vertex.
%In practice, the tricky part is to set $t$ in a non-asymptotic regime.
Given $k$, the optimal kernel width $t$ is found by searching for the minimizer of \eqref{eq:mean equivariance error}.
\Figref{kernel_widths} shows the optimal kernel widths found for various resolutions of the HEALPix sampling.
As predicted by the theory, $t_n \propto n^\beta, \beta \in \R$.
Importantly however, the optimal $t$ also depends on the number of neighbors $k$.

Considering the HEALPix sampling, \citet{perraudin2019deepspherecosmo} connected each vertex to their 8 adjacent vertices, computed the weights with \eqnref{weights:belkin}, and heuristically set $t$ to half the average squared Euclidean distance between connected vertices.
This heuristic however seems to over-estimate $t$ (see \figref{kernel_widths}).
% We observe that using a small kernel size $t$, \cite{perraudin2019deepspherecosmo} would probably be more equivariant.

@@

\paragraph{Theoretical convergence.}
\begin{wrapfigure}{r}{5cm}
	\centering
	\includegraphics[width=\linewidth]{patch}
	\caption{The patch around the $i$th sampling point.}
\end{wrapfigure}
We extend the work of \citep{belkin2005towards} to a sufficiently regular, deterministic sampling.
Following their setting, we will work with the extended graph Laplacian operator as the linear operator $L_n^t: L^{2}(\mathcal{M}) \rightarrow L^{2}(\mathcal{M})$ such that
\begin{equation} \label{eq:Heat Kernel Graph Laplacian operator}
	 L_n^t f(y) := \frac{1}{n}\sum_{i=1}^{n} e^{ -\frac{\|x_i-y\|^2}{4t}} \left(f(y)-f(x_i)\right).
\end{equation}
This operator extends the graph Laplacian operator previously defined with the weighting scheme \eqnref{weights:belkin} to each point of the the sphere.
% \mdeff{Can we avoid to introduce general manifolds and stay on the sphere?} \nati{yes}
As the radius of the kernel $t$ will be adapted to the number of samples, we also rescale the graph Laplacian operator as $\hat{L}_n^t := \frac{|\mathbb S^2|}{4\pi t^2}L_n^t$.
Hence, we have $\b{L}_n^t \b{f} = T_\V L_n^t f$.
% $\b{\hat L} = \frac{|\mathbb S^2|}{4\pi t^2}\b L_n^t$. \nati{why define this, we are not using it}
Given a sampling $\V$, we define $\sigma_i$ to be the patch of the surface of the sphere corresponding to $x_i$, $A_i$ its corresponding area, and $d_i$ the largest distance between the center $x_i$ and any point on the surface $\sigma_i$.
Define $d^{(n)} := \max_{i=1, \dots, n} d_i$ and $A^{(n)} := \max_{i=1, \dots, n} A_i$.
%Given any rotation $g\in SO(3)$, we define the corresponding rotation operator $\Rg$ such that for any function $f$ on the sphere
%\mdeff{Already defined in problem formulation.}
%\begin{equation} \label{eq:rotation operator}
%	\Rg f(x) := f(g^{-1}(x)).
%\end{equation}
\begin{theorem}
	For a sampling $\V$ of the sphere that is equi-area and such that $d^{(n)} \leq \frac{C}{n^\alpha}, \ \alpha\in (0,\linefrac{1}{2}]$, for all $f: \S^2 \rightarrow \R$ Lipschitz with respect to the Euclidean distance in $\R^3$, for all $y\in\S^2$, there exists a sequence $t_n = n^\beta$, $\beta\in\mathbb R$ such that
% 	the rescaled Heat Kernel Graph Laplacian $\frac{|\S^2|}{4\pi t_n^2}\b {L}^t_n$ converges pointwise to the Laplace Beltrami operator on the sphere $\Delta{\S^2}$
	\begin{equation*}
		\lim_{n\to\infty}\hat{L}_n^{t_n}f(y) = \triangle_{\S^2}f(y).
	\end{equation*}
	\label{theo:pointwise convergence for a regular sampling}
\end{theorem}
As a direct implication of this theorem, the rescaled graph Laplacian (the discrete counterpart of $\hat L^t_n$) converges pointwise towards the Laplace-Beltrami operator $\Delta_{\S^2}$ in the sense explained in \citep{belkin2007convergence}.
This is a major step towards equivariance as the Laplace-Beltrami operator does commute with rotation. Based on this property, we show that the graph Laplacian is equivariant to rotation in the limit of infinite sampling.

@@

% 3.1) explain what the figure tells us
For $k=8$ neighbors, selecting the optimal value of $t$ improves on \citet{perraudin2019deepspherecosmo}, whose graph also was constructed to have at most 8 neighbors.
The equivariance error further decreases by connecting more neighbors,  at the expense of computations.
Increasing the resolution decreases the equivariance error in the high frequencies.
% Increasing the graph resolution of the graph of \citet{khasanova2017sphericalcnn} does not seem to improve its equivariance.
Moreover, we observe that the weighting scheme \eqnref{weights:khasanova} does indeed not lead to convolutions that are equivariant to all rotations $g \in SO(3)$.
We however verified that the convolutions are equivariant to longitudinal and latitudinal rotations, as intended.

\begin{figure}
	\begin{minipage}{0.6\linewidth}
		\centering
		\includegraphics[width=1\linewidth]{all_in_one.pdf}
		\caption{
			Mean equivariance error \eqref{eq:mean equivariance error}.
			The number of pixels is $n = 4 s^2$ for the equiangular sampling and $n = 12 s^2$ for the HEALPix.
		}
		\label{fig:equivariance_error}
	\end{minipage}
	\hfill
	\begin{minipage}{0.35\linewidth}
		\centering
		\includegraphics[width=\linewidth]{kernel_widths}
		\caption{Kernel widths.}
		\label{fig:kernel_widths}
		\vspace{1em}
		\includegraphics[height=4.5em]{lamp_000018}
		\hfill
		\includegraphics[height=4.5em]{lamp_000018_sphere_nobar}
		\caption{3D object represented as a spherical depth map.}
		\label{fig:depthmap}
		\vspace{1em}
		\includegraphics[width=\linewidth]{spectrum}
		\caption{Power spectral densities.}
		\label{fig:spectrum}
	\end{minipage}
% \vspace{-0.5cm}
\end{figure}

@@

Theorem \ref{theo:pointwise convergence for a regular sampling}, is then an immediate consequence of Proposition \ref{prop:2} and \ref{prop:3}.
\begin{proof}[Proof of Theorem \ref{theo:pointwise convergence for a regular sampling}]
	Thanks to Proposition \ref{prop:2} and Proposition \ref{prop:3}	we conclude that $\forall y\in\S^2 $
	\begin{equation*}
	\lim_{n\to\infty}\frac{1}{4\pi t_n^2} L_n^{t_n}f(y) =  \lim_{n\to\infty}\frac{1}{4\pi t_n^2} L^{t_n}f(y) = \frac{1}{|\S^2|}\triangle_{\S^2}f(y)
	\end{equation*}
\end{proof}

>>>>

% 1) summarize the work of Khasanova
Considering the equiangular sampling and graphs where each vertex is connected to 4 neighbors (north, south, east, west), \citet{khasanova2017sphericalcnn} designed a weighting scheme to minimize \eqref{eq:mean equivariance error} for longitudinal and latitudinal rotations\footnote{Equivariance to longitudinal rotation is essentially given by the equiangular sampling.}.
% because longitudinal rotation can be discrete
Their solution gives weights inversely proportional to Euclidean distances:
\begin{equation} \label{eqn:weights:khasanova}
	w_{ij} = \frac{1}{\norm{x_i-x_j}}.
\end{equation}
%leading to a high mean equivariance error if the mean is computed over rotations in SO(3).
While the resulting convolution is not equivariant to the whole of $SO(3)$ (\figref{equivariance_error}), it is enough for omnidirectional imaging because, as gravity consistently orients the sphere, objects only rotate longitudinally or latitudinally.

% 2) explain our approach
To achieve equivariance to all rotations, we take inspiration from \citet{belkin2005towards}.
They prove that for a \emph{random uniform sampling}, the graph Laplacian $\b{L}$ built from weights
\begin{equation} \label{eqn:weights:belkin}
	w_{ij} = e^{- \frac{1}{4t} \norm{x_i - x_j}^2}
\end{equation}
converges to the Laplace-Beltrami operator $\Delta_{\mathbb{S}^2}$ as the number of samples grows to infinity.
This result is a good starting point as $\Delta_{\mathbb{S}^2}$ commutes with rotation, i.e., $\Delta_{\mathbb{S}^2}\Rg = \Rg\Delta_{\mathbb{S}^2}$.
% and (ii) is diagonalized by the spherical harmonics.
While the weighting scheme is full (i.e., every vertex is connected to every other vertex), most weights are small due to the exponential.
We hence make an approximation to limit the cost of the convolution \eqnref{graph_conv} by only considering the $k$ nearest neighbors ($k$-NN) of each vertex.
%In practice, the tricky part is to set $t$ in a non-asymptotic regime.
Given $k$, the optimal kernel width $t$ is found by searching for the minimizer of \eqref{eq:mean equivariance error}.
\Figref{kernel_widths} shows the optimal kernel widths found for various resolutions of the HEALPix sampling.
As predicted by the theory, $t_n \propto n^\beta, \beta \in \R$.
Importantly however, the optimal $t$ also depends on the number of neighbors $k$.

Considering the HEALPix sampling, \citet{perraudin2019deepspherecosmo} connected each vertex to their 8 adjacent vertices in the tiling of the sphere, computed the weights with \eqnref{weights:belkin}, and heuristically set $t$ to half the average squared Euclidean distance between connected vertices.
This heuristic however over-estimates $t$ (\figref{kernel_widths}) and leads to an increased equivariance error (\figref{equivariance_error}).
% We observe that using a small kernel size $t$, \cite{perraudin2019deepspherecosmo} would probably be more equivariant.

@@

\paragraph{Theoretical convergence.}
\begin{wrapfigure}{r}{5cm}
	\centering
	\includegraphics[width=\linewidth]{patch}
	\caption{The patch around the $i$th sampling point.}
\end{wrapfigure}
We extend the work of \citep{belkin2005towards} to a sufficiently regular, deterministic sampling.
Following their setting, we will work with the extended graph Laplacian operator as the linear operator $L_n^t: L^{2}(\mathcal{M}) \rightarrow L^{2}(\mathcal{M})$ such that
\begin{equation} \label{eq:Heat Kernel Graph Laplacian operator}
	 L_n^t f(y) := \frac{1}{n}\sum_{i=1}^{n} e^{ -\frac{\|x_i-y\|^2}{4t}} \left(f(y)-f(x_i)\right).
\end{equation}
This operator extends the graph Laplacian operator previously defined with the weighting scheme \eqnref{weights:belkin} to each point of the the sphere.
% \mdeff{Can we avoid to introduce general manifolds and stay on the sphere?} \nati{yes}
As the radius of the kernel $t$ will be adapted to the number of samples, we also rescale the graph Laplacian operator as $\hat{L}_n^t := \frac{|\mathbb S^2|}{4\pi t^2}L_n^t$.
Hence, we have $\b{L}_n^t \b{f} = T_\V L_n^t f$.
% $\b{\hat L} = \frac{|\mathbb S^2|}{4\pi t^2}\b L_n^t$. \nati{why define this, we are not using it}
Given a sampling $\V$, we define $\sigma_i$ to be the patch of the surface of the sphere corresponding to $x_i$, $A_i$ its corresponding area, and $d_i$ the largest distance between the center $x_i$ and any point on the surface $\sigma_i$.
Define $d^{(n)} := \max_{i=1, \dots, n} d_i$ and $A^{(n)} := \max_{i=1, \dots, n} A_i$.
%Given any rotation $g\in SO(3)$, we define the corresponding rotation operator $\Rg$ such that for any function $f$ on the sphere
%\mdeff{Already defined in problem formulation.}
%\begin{equation} \label{eq:rotation operator}
%	\Rg f(x) := f(g^{-1}(x)).
%\end{equation}
\begin{theorem}
	For a sampling $\V$ of the sphere that is equi-area and such that $d^{(n)} \leq \frac{C}{n^\alpha}, \ \alpha\in (0,\linefrac{1}{2}]$, for all $f: \S^2 \rightarrow \R$ Lipschitz with respect to the Euclidean distance in $\R^3$, for all $y\in\S^2$, there exists a sequence $t_n = n^\beta$, $\beta\in\mathbb R$ such that
% 	the rescaled Heat Kernel Graph Laplacian $\frac{|\S^2|}{4\pi t_n^2}\b {L}^t_n$ converges pointwise to the Laplace Beltrami operator on the sphere $\Delta{\S^2}$
	\begin{equation*}
		\lim_{n\to\infty}\hat{L}_n^{t_n}f(y) = \laplacian f(y).
	\end{equation*}
	\label{theo:pointwise convergence for a regular sampling}
\end{theorem}
As a direct implication of this theorem, the rescaled graph Laplacian (the discrete counterpart of $\hat L^t_n$) converges pointwise towards the Laplace-Beltrami operator $\Delta_{\S^2}$ in the sense explained in \citep{belkin2007convergence}.
This is a major step towards equivariance as the Laplace-Beltrami operator does commute with rotation. Based on this property, we show that the graph Laplacian is equivariant to rotation in the limit of infinite sampling.

@@

% 3.1) explain what the figure tells us
\Figref{equivariance_error} shows that the weighting scheme \eqnref{weights:khasanova} from \citep{khasanova2017sphericalcnn} does indeed not lead to a convolution that is equivariant to all rotations $g \in SO(3)$.\footnote{We however verified that the convolution is equivariant to longitudinal and latitudinal rotations, as intended.}
For $k=8$ neighbors, selecting the optimal kernel width $t$ improves on \citep{perraudin2019deepspherecosmo} at no cost, highlighting the importance of this parameter.
Increasing the resolution decreases the equivariance error in the high frequencies, an effect most probably due to the sampling.
Most importantly, the equivariance error decreases when connecting more neighbors.
Hence, the number of neighbors $k$ gives us a precise control of the tradeoff between cost and equivariance.

\begin{figure}
	\begin{minipage}{0.6\linewidth}
		\centering
		\includegraphics[width=\linewidth]{equivariance_error}
		\caption{
			Mean equivariance error \eqref{eq:mean equivariance error}.
			There is a clear tradeoff between equivariance and computational cost, governed by the number of vertices $n$ and edges $kn$.
		}
		\label{fig:equivariance_error}
	\end{minipage}
	\hfill
	\begin{minipage}{0.35\linewidth}
		\centering
		\includegraphics[width=\linewidth]{kernel_widths}
		\caption{Kernel widths.}
		\label{fig:kernel_widths}
		\vspace{1em}
		\includegraphics[height=4.5em]{lamp_000018}
		\hfill
		\includegraphics[height=4.5em]{lamp_000018_sphere_nobar}
		\caption{3D object represented as a spherical depth map.}
		\label{fig:depthmap}
		\vspace{1em}
		\includegraphics[width=\linewidth]{spectrum}
		\caption{Power spectral densities.}
		\label{fig:spectrum}
	\end{minipage}
% \vspace{-0.5cm}
\end{figure}

@@

Theorem \ref{theo:pointwise convergence for a regular sampling}, is then an immediate consequence of Proposition \ref{prop:2} and \ref{prop:3}.
\begin{proof}[Proof of Theorem \ref{theo:pointwise convergence for a regular sampling}]
	Thanks to Proposition \ref{prop:2} and Proposition \ref{prop:3}	we conclude that $\forall y\in\S^2 $
	\begin{equation*}
	\lim_{n\to\infty}\frac{1}{4\pi t_n^2} L_n^{t_n}f(y) =  \lim_{n\to\infty}\frac{1}{4\pi t_n^2} L^{t_n}f(y) = \frac{1}{|\S^2|}\laplacian f(y)
	\end{equation*}
\end{proof}
